Perceptual Symbol Grounding in \\Holographic Memories

### 2016-12-12 papers/hrr-paper/paper.tex
Rewrote, corrected and rearranged Abstract, Introduction and sections I and II

\begin{document}
	%
	% paper title
	% Titles are generally capitalized except for words such as a, an, and, as,
	% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
	% not capitalized unless they are the first or last word of the title.
	% Linebreaks \\ can be used within to get better formatting as desired.
	% Do not put math or special symbols in the title.
	\title{Perceptual symbol grounding in \\Holographic memories}

	% author names and affiliations
	% use a multiple column layout for up to three different
	% affiliations
	\author{\IEEEauthorblockN{Igor Peric, Stefan Ulbrich, Alexandru Lesi, Daniel Spies, R{\"u}diger Dillman, Marius Z{\"o}llner}
		\IEEEauthorblockA{FZI Forschungszentrum Informatik\\
			Karlsruhe, Germany\\
			Email: $\{$peric, ulbrich, lesi, spies, dillmann, zoellner$\}$@fzi.de}
	}

@@

	% As a general rule, do not put math, special symbols or citations
	% in the abstract
	\begin{abstract}
		Vector Symbolic Architectures (VSA) define a set of operations for association, (compressed) storage, manipulation and retrieval of symbolic concepts, represented as fixed-length vectors from ${\rm I\!R}^N$.
A specific instance of VSA, Holographic Reduced Representations (HRR), has proven to exhibit properties similar to human short-term memory and is, as such, interesting for computational modeling.

		In this paper we extend HRR approach by introducing implicit, topology-preserving, encoding/decoding procedure using radial basis functions.
Extended approach involves randomly chosen permutation (for encoding) and its inverse (for decoding), necessary for uniform distribution of signal in Fourier space in which embedding takes place.
These novel encoding schemes, together with extended operation pipeline, play the role of so called clean-up modules after memory retrieval (e.g. self-organizing maps).
Effectively, encoding implicitly represents the symbol, so no further lookup is needed.

		We show that newly introduced encoding scheme didn't have negative impact on memory capacity in comparison with original author's capacity benchmark.
We also test our memories in three different robotics tasks: mapping and recall of an area by a mobile agent, top-down visual attention shift and state machine scripting (holographic controllers).
	\end{abstract}

@@

	In addition, every VSA can be considered as a lossy compression technique, since binary operations on N-element vectors return N-element vectors as well.

	Depending of the way these three operations are implemented we can distinguish several popular VSAs, such as Holographic Reduced Representations \cite{Plate:1995:HolographicReducedRepresentations}, Circular HRRs \cite{DeVine:2010:Semanticoscillations}, Binary Spatter Codes \cite{Kanerva:1994:SpatterCodeEncoding}, Holographic Graph Neurons \cite{Kleyko:2016:HolographicGraphNeuron}, Geometric Algebras \cite{Patyk-Lonska:2011:DistributedRepresentationsBased} and Random Permutations \cite{Recchia:2015:EncodingSequentialInformation}.
The most widely used ones are HRRs and they are briefly explained in section \ref{sec:hrr}.
Authors in \cite{Golosio:2015:CognitiveNeuralArchitecture} have used them to solve word embedding problem for extremely large dataset by replacing computationally expensive big matrix factorization with iterative estimation algorithm based on HRRs.
Some of the other work is centered around pattern recognition \cite{Kleyko:2016:PatternRecognitionVector}, functional large-scale cognitive architecture models \cite{Eliasmith:2012:LargeScaleModel} and representing scene structure \cite{scene-structure}.
Some preliminary theoretical assumptions have been made about ability to use them as agent controllers \cite{Levy:2013:LearningBehaviorHierarchies}, where a symbolic program is represented with a set of combined symbolic actions associated with actuator symbols and symbolic sensing.

@@

	Authors in \cite{Eliasmith:2012:LargeScaleModel} have demonstrated functional cognitive model Spaun, implemented as HRRs using biologically plausible models of neuron dynamics, is capable of performing 7 different cognitive tasks.
In serial memory task they show that model of forgetting exhibits similar behavior as human working memory due to limited capacity.
Besides that, they report similarity between human working memory capacity and average number of items (symbols) that their model is capable of holding in memory.
This successful model of forgetting is one of the reasons our work is centered around HRRs.

	Second reason is the fact that HRRs are the only VSAs up to date operating on convolution operations in their core.
This is appealing if we take into consideration numerous recent advancement in deep learning and convolutional neural networks.
There are two basic reasons for success of these networks: training on big data and using convolution as the main architectural constraint.
Although description of biophysical mechanism underlying convolution and concept of weight sharing is still missing, it is important to reuse already proven concepts (such as convolutional filtering) in as general way as possible.
In other words, addressing multiple fundamentally different problems with the same or similar computational principles can be considered as a step towards unified brain processing theory.

@@

	\textbf{Association} (also called "binding") is an operation of linking two operands together, resulting in a vector of the same size which is dissimilar (orthogonal) to both original vectors, but contains scrambled information from both of the operands.
For instance, linking the property "red" to object "triangle" can be done using binding operation:

@@

	\textbf{Superposition} is operation of extending a set of already memorized items with another item (or binding of two or more of them).
As an example, a memory of "red triangle and blue circle" can be formed in following way:
	\begin{equation}
	memory_2 = memory_1 + blue \otimes circle
	\end{equation}

	\textbf{Probing} is operation of extracting a noisy version of associated item from a joint set memory.
I.e. a question "What is the color of triangle?" could be asked like this:

@@

	Probing operation assumes existence of inverse element to association, so previous equation can be interpreted and explained in following way:
	\begin{multline}
	memory_2 \otimes triangle'=\\
	(red \otimes triangle + blue \otimes circle) \otimes triangle'=\\
	red \otimes triangle  \otimes triangle' + blue \otimes circle \otimes triangle' =\\
	red + blue \otimes circle \otimes triangle'  = red + noise
	\approx red
	\end{multline}

	where $triangle'$ represents inverse of $triangle$.
Resulting vector will be composition of something that is very similar to $red$ and something that doesn't resemble anything meaningful from symbol table, so it can be treated as neglectable noise.

@@

	Process of converting noisy result of probing into the closest know symbol is called \textbf{clean-up procedure}.
Straightforward way of implementing clean-up memory is using a lookup table, while other researchers show that any type of attractor network can be used (e.g. self-organizing maps).

@@

	Holographic Reduced Representations are VSAs which implement operation of binding as circular convolution:

@@

	where $z$ is the result of convolution, $x$ and $y$ are symbolic vectors of size $N$.
Superposition is implemented as simple element-wise addition.
Probing is implemented using inverse of circular convolution (circular correlation):

@@

	In order to use various types of inputs with HRRs it is necessary to encode them in vectors of fixed size.
	The basic approach is to pick these vectors randomly once for each symbol and maintain these pairs in a separate data structure.
	This will consequently increase space complexity and may become infeasible, depending on the type of input used. For example, handling scalar symbolic encoding can be done by discretization of the whole input range into bins and assigning a random vector to each discrete value. In order to decode a specific symbolic vector, one must find maximum similarity element from symbol table, which requires O(M), where M is the number of symbols known to the system.

	Usual way of computing similarity is cosine distance, which for two vectors \pmb x and \pmb y is calculated as follows:

	This approach works quite well for object identifiers (names), but has some non-desirable properties when used on scalar encoding. These will be discussed in the following section.

@@

	\section{Experiments}

	\subsection{Visual short term memory and top-down attention model}

>>>>

\begin{document}
	%
	% paper title
	% Titles are generally capitalized except for words such as a, an, and, as,
	% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
	% not capitalized unless they are the first or last word of the title.
	% Linebreaks \\ can be used within to get better formatting as desired.
	% Do not put math or special symbols in the title.
	\title{Perceptual Symbol Grounding in \\Holographic Memories}

	% author names and affiliations
	% use a multiple column layout for up to three different
	% affiliations
	\author{\IEEEauthorblockN{Igor Peric, Alexandru Lesi, Daniel Spies, Stefan Ulbrich, R{\"u}diger Dillman, Marius Z{\"o}llner}
		\IEEEauthorblockA{FZI Forschungszentrum Informatik\\
			Karlsruhe, Germany\\
			Email: $\{$peric, lesi, spies, ulbrich, dillmann, zoellner$\}$@fzi.de}
	}

@@

	% As a general rule, do not put math, special symbols or citations
	% in the abstract
	\begin{abstract}
		Vector Symbolic Architectures (VSA) define a set of operations for association, (compressed) storage, manipulation and retrieval of symbolic concepts, represented as fixed-length vectors from ${\rm I\!R}^N$.
A specific instance of VSA, Holographic Reduced Representations (HRR), has proven to exhibit properties similar to human short-term memory and is, as such, interesting for computational modelling.

		In this paper we extend the HRR approach by introducing implicit, topology-preserving, encoding/decoding procedure using radial basis functions.
Extended approach involves randomly chosen permutation (for encoding) and its inverse (for decoding), necessary for uniform distribution of signal in Fourier space in which embedding takes place.
These novel encoding schemes, together with extended operation pipeline, play the role of so called clean-up modules after memory retrieval (e.g. self-organizing maps).
Effectively, encoding implicitly represents the symbol, so no further lookup is needed.

		We show that this newly introduced encoding scheme didn't have a negative impact on memory capacity in comparison with the original author's capacity benchmark.
We also test our memories in three different robotics tasks: mapping and recall of an area by a mobile agent, top-down visual attention shift and state machine scripting (holographic controllers). 
	\end{abstract}

@@

	In addition, every VSA can be considered as a lossy compression technique, since binary operations on N-element vectors return N-element vectors as well. These operations will be described in further detail in the following section.

	Depending on the way these three operations are implemented we can distinguish several popular VSAs, such as Holographic Reduced Representations \cite{Plate:1995:HolographicReducedRepresentations}, Circular HRRs \cite{DeVine:2010:Semanticoscillations}, Binary Spatter Codes \cite{Kanerva:1994:SpatterCodeEncoding}, Holographic Graph Neurons \cite{Kleyko:2016:HolographicGraphNeuron}, Geometric Algebras \cite{Patyk-Lonska:2011:DistributedRepresentationsBased} and Random Permutations \cite{Recchia:2015:EncodingSequentialInformation}.
The most widely used ones are HRRs and they are briefly explained in section \ref{sec:hrr}.
Authors in \cite{Golosio:2015:CognitiveNeuralArchitecture} have used them to solve word embedding problem for extremely large dataset by replacing computationally expensive big matrix factorization with an iterative estimation algorithm based on HRRs.
Some of the other work is centered around pattern recognition \cite{Kleyko:2016:PatternRecognitionVector}, functional large-scale cognitive architecture models \cite{Eliasmith:2012:LargeScaleModel} and representing scene structure \cite{scene-structure}.
Preliminary theoretical assumptions have been made about the ability to use them as agent controllers \cite{Levy:2013:LearningBehaviorHierarchies}, where a symbolic program is represented with a set of combined symbolic actions associated with actuator symbols and symbolic sensing.

@@

	Authors in \cite{Eliasmith:2012:LargeScaleModel} have demonstrated that the functional cognitive model Spaun is capable of performing 7 different cognitive tasks when implemented as HRRs using biologically plausible models of neuron dynamics.
They further show that in serial working memory tasks the model of forgetting exhibits similar behaviour as human working memory due to limited capacity.
Lastly they report similarity between the human working memory capacity and the average number of items (symbols) that their model is capable of holding in memory.

	This successful model of forgetting is one of the reasons our work is centred around HRRs. Another reason is the fact that HRRs are the only VSAs up to date operating on convolution operations in their core.
This is appealing if we take the numerous recent advancement in deep learning and convolutional neural networks into consideration.
These networks succeed due to training on big data and using convolution as the main architectural constraint.
Although the description of the biophysical mechanism underlying convolution and concept of weight sharing are still missing, it is important to reuse proven concepts, such as convolutional filtering, in as general way as possible.
In other words, addressing multiple fundamentally different problems with similar computational principles can be considered a step towards a unified brain processing theory.

@@

In order to use VSA operations all types of required inputs need to first be represented as symbols, which are vectors of fixed size. The basic approach is to pick these vectors randomly once for each input used in any operation and maintain these pairs in a separate data structure.	
	\\

	\textbf{Association} (also called "binding") is an operation of linking two operands together, resulting in a vector of the same size which is dissimilar (or orthogonal) to the original vectors, but also contains scrambled information from both of them.
For instance, linking the property "red" to object "triangle" can be done using the binding operation as follows:

@@

	\textbf{Superposition} is the operation of extending a set of items that have already been memorized with another item (or binding of two or more of them).
As an example, a memory of "red triangle and blue circle" can be formed in following way:
	\begin{equation}
	memory_2 = memory_1 + blue \otimes circle
	\end{equation}

	\textbf{Probing} is the operation of extracting a noisy version of an associated item from a joint set memory.
I.e. in this scenario the question "What color is the triangle?" could be asked like this:

@@

	The probing operation assumes the existence of an inverse element to association, so the previous equation can be interpreted and explained in following way:
	\begin{multline}
	memory_2 \otimes triangle'=\\
	(red \otimes triangle + blue \otimes circle) \otimes triangle'=\\
	red \otimes triangle  \otimes triangle' + blue \otimes circle \otimes triangle' =\\
	red + blue \otimes circle \otimes triangle'  = red + noise
	\approx red
	\end{multline}

	where $triangle'$ represents inverse of $triangle$.
The resulting vector will be a composition of something that is very similar to $red$ and something that doesn't resemble anything meaningful from the symbol table, so it can be treated as negligible noise.	A usual way of computing similarity is the cosine distance, which for two vectors \pmb x and \pmb y is calculated as follows:

@@

	Obtained values of similarity clearly distinguish the matching vectors. Concrete values that highlight this will be later presented in section \ref{sec:experiments}.

	The process of converting a noisy result of probing into the closest known symbol is called the \textbf{clean-up procedure}.
A straightforward way of implementing clean-up memory is using a lookup table, while other researchers show that any type of attractor network can be used (e.g. self-organizing maps).

	In the case of look-up tables, in order to decode a specific symbolic vector, the elements from a symbol table are iterated over, the similarity to each of the entries is calculated and the symbol which presented maximum similarity is returned. A threshold for the similarity value is also used in order to exclude false-positives. This look-up requires O(M), where M is the number of symbols known to the system. This approach works quite well for object identifiers (names), but has some non-desirable properties when used on scalar encoding.
	//TODO Lookup Example

	This will consequently increase space complexity and may become infeasible, depending on the type of input used. While looking for existing elements may not require the entire table to be checked probing for non-existent ones will always traverse all entries only to show up empty at the end.When handling scalars symbolic encoding can be done by discretization of the whole input range into bins and assigning a random vector to each discrete value. This however may often prove to not satisfy the task needs any more.

@@

	Holographic Reduced Representations are VSAs which implement the operation of binding as circular convolution:

@@

	where $z$ is the result of convolution, and $x$ and $y$ are symbolic vectors of size $N$.
Superposition is implemented as a simple element-wise vector addition.
Probing is implemented using inverse of circular convolution (circular correlation):

@@

@@

	\section{Experiments}
	\label{sec:experiments}
	\subsection{Visual short term memory and top-down attention model}
